### 2015-01-24 algorithms.tex

\section{Algorithm}
\label{sec:background}
\input{algorithm/introduction.tex}
\input{algorithm/filtering.tex}
\input{algorithm/salience.tex}
\input{algorithm/ap.tex}
\input{algorithm/novelty.tex}