### 2015-01-25 algorithm/novelty.tex
Read and commented on algorithm sections

\subsubsection{Ensuring Novelty with Respect to Previous Updates}
\label{sec:novelty}
After we run AP clustering, we have a set of candidates for the current hour.  In order to avoid
%KM - word change
%outputting
producing
redundant updates, we computed the semantic similarity,
%KM - You could give the semantic similarity algorithm a name. THat woudl make it easier to refer to. In any case, you need to specify the method you use here. I added clause below. Coudl remove it if you name the algorithm.
using the same method as used in salience prediction,
between the exemplar sentences and previously output sentences.  We then removed those sentences which were above some similarity threshold.  We selected the threshold by inspecting the distribution of similarities between sentences associated with the same sub-event in the training set and the distribution of similarities between random pairs of sentences.

% and preferences. When the clustering has converged we emit all exemplar
% sentences (cluster centers) whose preference is $\lambda$ standard deviations
% above the mean preference for the current time period and who do not belong
% to singleton clusters as updates. These updates and their preferences are
% retained for penalty calculations in subsequent time steps.
%