### 2015-02-22 introduction.tex
Moving to acl style

\label{sec:introduction}
During crises, information is critical for first responders,
 crisis management organizations, and those caught in the event.  
When the event is significant, as in the case of Hurricane
Sandy, the amount of content produced by traditional news outlets,
government agencies, relief organizations, and social media can vastly
overwhelm those trying to monitor the situation. 
Crisis informatics\cite{?} is dedicated to finding methods for sharing the
right information in a timely fashion with relief organizations during a major crisis. 
Research in this field has focused on
human-in-the-loop approaches ranging from on the ground information gathering to
crowdsourced reporting and disaster management.

Multi-document summarization offers potential for enabling automatic updates of
relevant, salient information at regular intervals. It would provide
information even when human volunteers are unable to and would filter out
unnecessary and irrelevant detail. 
While methods for identifying,
tracking, and summarizing events from text based input have been explored
extensively
%KM - I think we should consider adding other event summarizers. 
(e.g.,
\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}), 
these experiments were not developed to handle streaming data from the large
and heterogeneous environment of the modern web. 
These methods also rely heavily on redundancy which is suboptimal for time
sensitive domains where there is a high cost in delaying information.

%\cite{?} to enable people on the ground to update
%crisis interfaces with information about needs. Others have used crowd
%sourcing \cite{?},

%At times of crisis, however, social media can overwhelm current information
%systems with the quantity of information, much of which is irrelevant,
%unnecessary detail, or out of date.

%Neither were they developed to address the
%information needs that arise during  crisis
%situations. 
%KM - this last sentence sounds somewhat negative.
%there is still a need for robust and scalable 
%methods for automatic summarization.

In this paper, we present an update summarization system to track events
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster, and integrates these predictions into
a clustering based multi-document summarization system. 
Our system is better able to adapt to
dynamic changes in input volume which adversely 
effect methods of summarization that use either pairwise
comparisons as a proxy for salience or predict salience but
ignore redundancy in the input alone.

%We train a regression
%model to predict sentence salience and use these predictions to bias the
%formation of sentence clusters around more salient regions in the input space
%using affinity propagation (AP) clustering.  We adapt AP to incorporate
%the salience predictions
%as well as pairwise similarities among input sentences to identify
%\emph{exemplar} sentences, which we use as our summary output.

%ignoring features of importance that are intrinsic to the
%sentences themselves.

%KM - Chris - add a short overview on the kinds of things we measure and
%results so people are encouraged to read ahead. Keep it general.

In addition to the tight integration between clustering and salience
prediction, our approach also exploits knowledge about disaster to determine
salience. Thus, salience does not just represent importance within a set of
documents; it also represents both how typical this sentence is of the input 
event
type (i.e., disaster, hurricane, tornado) and whether it specifies information
about this particular disaster. 
Our feature representation includes a set of language models, one for each
disaster type, to measure typicality of the sentence for the current event 
type, the distance of mentioned locations from the center of
the disaster, and the change in word frequencies over the time of the event.
% to represent its likelihood of referring to the input disaster. 
%We also utilize the change in word frequencies as we process the stream 
%to measure the changing importance of updates over time.

Our approach achieves suprerior ROUGE scores compared to multiple baselines.
Additionally, we introduce a novel method for determining the 
coverage of atomic facts in a summary, and show our system's
superiority on this metric.

The remainder of the paper is organized as follows.
We begin with a review of related work
in the information retrieval and multi-document
summarization literature. (Rest of outline goes here...)

%Our main contributions are three fold. First, we present a straightforward 
%framework for 
%combining prior beliefs with a clustering algorithm. Second, we demonstrate
%the empirical effectiveness of this approach on the update summarization
%task. Finally, we present a novel feature represention for sentence salience
%in the crisis/disaster domain.