### 2015-02-22 motivation.tex

%KM If we are short on space, not sure how much this section is needed as it's
%somewhat repetitive of intro.

%KM - Chris, could you get some good references to work in the crisis
%informatics space.
Crisis informatics\cite{?} is dedicated to finding methods for sharing the
right information in a timely fashion with relief organizations during a major crisis. With the
increasing impact of climate change, the world is seeing an increase in
disasters such as hurricanes, tornadoes, flooding and typhoons, all of which
cause major damage and wreak havoc with food supplies, housing, and health
issues. Health epidemics, such as the ebola crisis in West Africa, also create
a need for timely information about where problems are greatest. Social and
political 
crises, such as the current situation in Syria, create similar needs for
humanitarian assistance.

At times of crisis, however, social media can overwhelm current information
systems with the quantity of information, much of which is irrelevant,
unnecessary detail, or out of date. Many approaches have focused on
human-in-the-loop approaches \cite{?} to enable people on the ground to update
crisis interfaces with information about needs. Others have ..
%KM - Chris - fill this out a bit.

%KM - PRobably should make a little more inclusive, there are graph based
%models that use methods such lexical similarity to build the graph,
%probabilistic methods (e.g., based on word probabilities, regression or
%topical signatures). There are also methods that use topic segmentation or
%modeling and select sentences for each topic.

Multi-document summarization offers potential for enabling automatic updates of
relevant, salient information at regular intervals. It would provide
information even when human volunteers are unable to and would filter out
unnecessary and irrelevant detail.

We conceptualize the space of existing summarization approaches as falling into
one of three categories, each with specific tradeoffs with respect to update 
summarization. First, centrality-focused approaches (including graph,
cluster, and centroid methods) are very natural for retrospective analysis in
the sense that they let the data "speak for itself." However, they rely
chiefly on
redundancy. When applied to an unfolding event, there may not exist enough
redundant content at the event onset for these methods to exploit.
Once the event onset has passed, however, the redundancy reduction of these 
methods is quite beneficial.

The second category, predictive approaches,
includes ranking and classification based methods.
These methods do not control for redundancy as well as the first group, 
and will most likely be overwhelmed by web-scale summarization.
On the other hand, these methods are not dependent on redundancy.

The last category includes probabilistic, information theoritic, and set cover
approaches. While these methods are generally focused on producing diverse
summaries, they are difficult to adapt to the streaming setting, where 
we do not necessarily have a fixed summary length and the corpus to be
summarized contains many irrelevent sentences, i.e. there are large
portions of the corpora that we specificly want to avoid.

Our main contribution is a framework for combining the strengths of the 
centrality-based methods and predictive methods using salience predictions (we 
prefer the term salience over rank) to bias cluster formation around high 
salience sentences.% by combining regression with AP clustering.
By algorithmically making the trade off between redundancy
signals and salience signals, we mitigate the effect of 
different input sizes. 
When input is small, individual characteristics
of the input determine centrality; when input is large, neighbor similarity 
becomes more important.

Second, we validate this method with the application to the update
summarization
task. We experiment on events of dramatically different scope and duration,
and demonstrate improved average performance on ROUGE. Even within events
the document streams can have widely varying levels of volume, corresponding
to the day/night cycle (see fig. ?? for an example).

Finally, we present an analysis of our feature representation for salience
disaster domain. While we focus on a particular domain, the feature templates
themselves are generally applicable to streaming/update summarization.

%Most generic multi-document summarization approaches involve either sentence 
%clustering or ranking sentences according to metrics that measure the salience
%of their words. 
%When the summarization task
%is broad or underspecified,
%these methods are appropriate as they let the data
%``to speak for itself.''
%When there are more constraints on the output, as in query-focused
%summarization or the crisis update task we present here,
%results are more accurate when
%appropriate features for the ranking approach are used (e.g., amount of word
%overlap for the query-focused approach).
%
%The primary contribution of this paper
%is a framework for combining both of these approaches using affinity
%propagation clustering as well as disaster-specific features.

%\begin{figure}
%\begin{tabular}{| p{8cm} |}
%\hline
%\textbf{Event: Hurricane Sandy}\\
%\includesvg[width=\columnwidth]{docfreq}\\
%\textbf{Example Nuggets:}\\
%$\cdot$ October 22nd Sandy strengthened from a tropical depression into a tropical storm\\
%$\cdot$ could turn towards Bermuda or go straight to the Eastern U.S.\\
%$\cdot$ 11 pm Oct 24 winds 50 mph\\
%%$\cdot$ 11 pm Oct 24 moving NNE at 9 knots\\
%$\cdot$ More than 1,000 people went to shelters in Jamaica\\
%\hline
%\end{tabular}
%\end{figure}