### 2015-02-26 experiments.tex
latest revision changes

We evaluate our system on two metrics: ROUGE \cite{?}, the most common automatic
summarization method and a modified method adapted 
from the Text Retrieval Conference Temporal Summarization (TREC-TS) track \cite{?}.

\subsection{Dataset}

Both evaluations make use of the TREC TS data from 2013 and 2014. 
This data includes 25 events and event metadata (e.g., a user
search query for the event, the event type, and event time frame).
Additionally, TREC assessors constructed a set of ground truth facts, or 
nuggets, for each event. The events took place from 20??-20?? and had 
an associated Wikipedia (WP) page. The WP pages were edited and updated 
concurrently with the events, so the diffs between revisions produce a similar
ouput to the ideal update summarization system. Assessors analyzed the diffs
and selected phrases that were important enough to include in an update 
summary.  
For the document stream, we use the news portion of the
 2014 TREC Stream Corpus.
The documents from this corpus come from hourly crawls of the web covering 
 ?, 20?? to ?, 20??. \ckcomment{Need to mention we further filtered the corpus}

\subsection{Training and Testing}

Of the 25 events in the TREC TS data, 24 are covered by the Stream Corpus.
From these 24, we set aside three events to use as a development set.
All system salience and similarity threshold parameters are tuned on the
development set.

We train a salience model for each event using 1000 sentences randomly sampled
from the event's document stream. Kernel length scale parameters  for each
feature group are fit to maximize
the model's log likelihood using ???.
At test time, we predict a sentence's salience using the average predictions
of the 23 other models---essentially a leave-one-out approach.

\subsection{ROUGE Evaluation}

ROUGE measures the ngram overlap between a model (i.e. human) summary 
and an automatically generated system summary. 
Model summaries for each event were constructed by concatenating the event's 
nuggets. 
See figure~\ref{fig:test} for an example model summary excerpt.
Generally, ROUGE evaluation assumes both model and system summaries
are of a bounded length. Since our systems are summarizing events over a span
of two weeks time, the total length of our system output is much longer than
the model. To address this, for each system/event pair, we sample with replacement
1000 random summaries of length less than or equal to the model summary 
(truncating the last sentence when neccessary). The final ROUGE score for the 
system are the average scores from these 1000 samples.

Because we are interested in system performance over time, we also evaluate 
systems at 12 hour intervals using the same regime as above. 
The model summaries in this case are retrospective, and so we can see how quickly
each system takes to achieve its maximum performance.

\subsection{Average Gain and Comprehensiveness}

We treat the event's nuggets as unique units of information.
When a system adds an update to its summary, it is potentially adding some
of this nugget information. It would be instructive to know how much unique
and novel information each update is adding on average to the summary.
To that end, we define

\[ \mathrm{Avg. Gain} = \frac{|S_n|}{|S|}% \;\;\; \mathrm{Recall} = \frac{|S_n|}{|N|}
\] 
where $S$ is the set of system updates, 
$S_n$ is the set of nuggets contained in $S$, and $|\cdot|$ is the number of
elements in the set.
To compute the set $S_n$ we match each system update to 0 or more nuggets, 
where an update matches a nugget if their semantic similarity is above 
a threshold. $S_n$ results from the unique set of nuggets matched.
Because an update can map to more than one nugget, it is possible to receive an
average gain greater than 1. An average gain of 1 would indicate that every
sentence was both relevant and contained a unique piece of information.

Additionally, we can use the nuggets to measure the completeness of an update 
summary. We define

\[ \mathrm{Comprehensiveness} = \frac{|S_n|}{|N|}\]

where $N$ is the set of event nuggets.
A comprehensiveness of 1 indicates that the summary has cover 
all nugget information for the event; the maximum
attainable comprehensiveness is 1.

Determining an optimal threshold to count matches is difficult so 
we evaluate at 
threshold values ranging
from .5 to 1, where values closer to 1 are more conservative estimates of
performance.
A qualitative assessment of matches 
suggests that semantic similarity values around .7 produce reasonable matches.
For example, the nugget 
\begin{blockquote}
watches and warnings in effect for much of the Mid-Atlantic states and 
southern New England 
\end{blockquote}
\noindent
matches with the update 
\begin{blockquote}
A tropical storm warning remains in effect for much of the coasts 
of North and South Carolina\end{blockquote}
\noindent
with a similarity of .7. 
The average semantic similarity of manual matches performed 
by TREC assessors was considerably lower at approximately .25, increasing
our confidence in the automatic matches in the .5-1 range.

An assessment of manual matchings done by TREC 
assessors revealed that the average cosine similarity

\subsection{Model Comparisons}

Our proposed model using AP+Salience is compared against several variants and baselines. All thresholds
are tuned on the development set.

\textbf{AP+Salience Time Ranked} : This model is everywhere the same as AP+Salience except at the update selection stage, 
where the algorithm selects exemplars in temporal order rather than salience order.

\textbf{AP} : In this model we use the AP clustering algorithm without the salience predictions.
The preferences are chosen as the median value of the ? matrix, as is commonly used with AP \cite{?}.
The update selection procedes in time order. As in the AP+Salience model we tune a similarity threshold,
and ignore exemplar sentences whose maximum similarity to a previous update is above the threshold.

\textbf{Rank by Salience} : In this model we rank the sentences by salience and select the updates at each hour that are
above a salience threshold and are below a similarity threshold (similar to the AP+Salience model).

\textbf{Hierarchichal Agglomerative Clustering (HAC)} : We also provide another clustering baseline, HAC using ?-linkage. 
This method was chosen over other clustering approaches because the number of clusters 
is not an explicit hyper parameter. To produce flat clusters from the hierarchical clustering, we flatten the HAC dendrogram
using the cophenetic distance criteria, i.e. observations in each flat cluster have no greater a cophenetic distance than a threshold.
Cluster centers are determined to be the sentence with highest cosine similariy to the flat cluster mean.
Cluster centers are examined in time order and are added to the summary if their similarity to previous updates is below a similarity threshold
(like the previous models).