### 2015-02-26 dataset.tex
Latest revision ck

For the document stream, we use the news portion of the
 2014 Trec KBA Stream Corpus.
The documents from this corpus come from hourly crawls of the web covering 
 ?, 20?? to ?, 20??. \ckcomment{Need to mention we further filtered the corpus}

Our experiments also make use of the Trec Temporal Summarization (TS) Track
 data from 2013 and 2014. 
This data includes 25 events and event metadata (e.g., a user
search query for the event, the event type, and event evaluation timeframe).  
All events occurred during the time span of the Trec KBA Stream Corpus.

Along with the metadata, NIST assessors constructed a set of ground truth nuggets for each event. 
Nuggets are brief and important text snippets that represent sub-events that should be conveyed
by an ideal update summary (see Figure~\ref{fig:nuggets} for examples).
In order
to accomplish this, for each event, assessors were provided with the
revision history of the Wikipedia page associated with the event.  
For example, 
the revision history for the Wikipedia page for `Hurricane Sandy' will 
contain text additions including those related to individual nuggets.  The assessment
task involves reviewing the Wikipedia revisions in the evaluation timeframe 
and marking the text additions capturing a new, unique nugget.  More detail
on this process can be found in the track description \cite{?}.