### 2015-02-27 conclusion.tex
Latest rev ck

In this paper, we have presented an update summarization system for the disaster domain,
and demonstrated improved system performance by integrating sentence salience with clustering.
Addtionally we have presented several interesting feature groups for use in disaster summarization.
In the future we would like to explore the applicability of this model and features
to more general domains.