### 2015-02-27 experiments.tex

We evaluate our system on two metrics: ROUGE \cite{?}, a common automatic
summarization method and an evaluation of system expected gain 
and comprehensiveness---metrics adapted 
from the Trec TS track \cite{?}.

\subsection{Training and Testing}

Of the 25 events in the TREC TS data, 24 are covered by the news portion 
of the Trec KBA Stream Corpus.
From these 24, we set aside three events to use as a development set.
All system salience and similarity threshold parameters are tuned on the
development set.

We train a salience model for each event using 1000 sentences randomly sampled
from the event's document stream. Kernel length scale parameters  for each
feature group are fit to maximize
the model's log likelihood using ???.

We perform a leave-one-out evaluation of each event.
At test time, we predict a sentence's salience using the average predictions
of the 23 other models.

\subsection{ROUGE Evaluation}

ROUGE measures the ngram overlap between a model (i.e. human) summary 
and an automatically generated system summary. 
Model summaries for each event were constructed by concatenating the event's 
nuggets. 
See figure~\ref{fig:test} for an example model summary excerpt.
Generally, ROUGE evaluation assumes both model and system summaries
are of a bounded length. Since our systems are summarizing events over a span
of two weeks time, the total length of our system output is much longer than
the model. To address this, for each system/event pair, we sample with replacement
1000 random summaries of length less than or equal to the model summary 
(truncating the last sentence when neccessary). The final ROUGE score for the 
system are the average scores from these 1000 samples.

Because we are interested in system performance over time, we also evaluate 
systems at 12 hour intervals using the same regime as above. 
The model summaries in this case are retrospective, 
and so we can see how quickly
each system is able to achieve its maximum performance.

\subsection{Expected Gain and Comprehensiveness}

NIST developed a series metrics for evaluating update summarization systems
as part of the TREC TS track.
We present results on two of these metrics, the expected gain and 
comprehensiveness.

\textbf{Expected Gain } We treat the event's nuggets as unique units of 
information.
When a system adds an update to its summary, it is potentially adding some
of this nugget information. It would be instructive to know how much unique
and novel information each update is adding on average to the summary.
To that end, we define

\[ \mathrm{\mathbb{E}[Gain]} = \frac{|S_n|}{|S|}% \;\;\; \mathrm{Recall} = \frac{|S_n|}{|N|}
\] 
where $S$ is the set of system updates, 
$S_n$ is the set of nuggets contained in $S$, and $|\cdot|$ is the number of
elements in the set.
To compute the set $S_n$ we match each system update to 0 or more nuggets, 
where an update matches a nugget if their semantic similarity is above 
a threshold. $S_n$ results from the unique set of nuggets matched.
Because an update can map to more than one nugget, it is possible to receive an
expected gain greater than 1.
%values greater than one can be thought of as a 
%compression ration, i.e. an expected gain of 10 indicates the average update
%contains 10 unique nuggets on average and that the summary is a 10~:~1 
%compression of the nugget information.
An expected gain of 1 would indicate that every
sentence was both relevant and contained a unique piece of information.

\textbf{Comprehensiveness } Additionally, we can use the nuggets to measure the completeness of an update 
summary. We define

\[ \mathrm{Comprehensiveness} = \frac{|S_n|}{|N|}\]

where $N$ is the set of event nuggets.
A comprehensiveness of 1 indicates that the summary has cover 
all nugget information for the event; the maximum
attainable comprehensiveness is 1.

Update-nugget matches are computed automatically, where a
match exists if the semantic similarity of the update/nugget pair is above 
a threshold.
Determining an optimal threshold to count matches is difficult so 
we evaluate at 
threshold values ranging
from .5 to 1, where values closer to 1 are more conservative estimates of
performance.
A manual inspection of matches 
suggests that semantic similarity values around .7 produce reasonable matches.
For example, the nugget 
\begin{blockquote}
watches and warnings in effect for much of the Mid-Atlantic states and 
southern New England 
\end{blockquote}
\noindent
matches with the update 
\begin{blockquote}
A tropical storm warning remains in effect for much of the coasts 
of North and South Carolina\end{blockquote}
\noindent
with a similarity of .7. 
The average semantic similarity of manual matches performed 
by NIST assessors was considerably lower at approximately .25, increasing
our confidence in the automatic matches in the .5--1 range.

\subsection{Model Comparisons}
We refer to our complete model as \textbf{AP+Salience}.  We compare this model again several variants and baselines intended to measure the contribution of different components. All thresholds
for all runs are tuned on the development set.

\textbf{Affinity Propagation only (AP) } The purpose of this model
is to directly measure the effect of integrating salience and clustering by providing a baseline that uses the identical 
clustering component, but without the salience information.
In this model, input sentences are apriori equally likely to be exemplars;
%clustering algorithm without the salience predictions.
the salience values are uniformly set as the median value of the 
input similarity scores, as is commonly used in the AP literature \cite{?}.
After clustering a sentence batch, the exemplars are examined in order
of increasing time since event start and selected as updates if their
maximum similarity to the previous updates is less than $\lambda_{sim}$,
as in the novelty filtering stage of AP+Salience.

\textbf{Hierarchichal Agglomerative Clustering (HAC)} : We provide another clustering baseline, single-linkage hierarchichal agglomerative clustering.
We include this baseline to show that AP+Salience is not just an improvement
over AP but centrality driven methods in general.
HAC was chosen over other clustering approaches because the number of clusters 
is not an explicit hyper-parameter. To produce flat clusters from the hierarchical clustering, we flatten the HAC dendrogram
using the cophenetic distance criteria, i.e. observations in each flat cluster have no greater a cophenetic distance than a threshold.
Cluster centers are determined to be the sentence with highest cosine similariy to the flat cluster mean.
Cluster centers are examined in time order and are added to the summary if their similarity to previous updates is below a similarity threshold $\lambda_{sim}$ as is done in the AP model.

\textbf{Rank by Salience (RS)} 
We also isolate the impact of our salience model in order to demonstrate 
the that the fusion of clustering and salience prediction improves over
predicting salience alone.
In this model we predict the salience of sentences as in step 1 for 
AP+Salience. %However, rather than clustering, we simply rank the sentences
%by salience.
We omit the clustering phase (step 2).
Updates are selected identically to step 3 of AP+Salience, proceeding 
in order of decreasing salience, selecting updates that are above a salience 
threshold $\lambda_{sal}$ and below a similarity threshold $\lambda_{sim}$
with respect to the previously selected updates.

%\fdcomment{consider removing AP+SalienceTR}
%\textbf{AP+Salience Time Ranked (AP+SalienceTR)} Our final model is identical 
%to the
%AP+Salience model except at the novelty filter stage.
%Within the novelty filtering stage, 
%AP+Salience inspects exemplars in order of decreasing salience when computing
%similarity to previous updates. Because the order in which we add updates
%affects future similarity calculations, we evaluate an alternate ordering 
%scheme (AP+SalienceTR) where exemplars are selected in order of increasing 
%time since event start. The purpose of this model is to give preference to 
%more recent updates, since timelyness is one of the desirable traits of our update
%summaries.
%