### 2015-02-27 feature_table.tex

\begin{figure}[t!]
\begin{tabular}{| l |} 
\hline
\textbf{Basic Features}\\
$\cdot$ Sent. position (normalized by doc. length) \\
$\cdot$ Sent. length \\
$\cdot$ Ratio of punc./caps./lower/upper to \\
$\;\;$ other chars. \\
%$\cdot$ Ratio of caps. to non-caps. chars. \\
%$\cdot$ Ratio of lowercase to other chars. \\
%$\cdot$ Ratio of uppercase to other chars. \\
$\cdot$ \# of caps. words (normalized by \# of words)\\
$\cdot$ \# of Person, Location, Org., etc.\\
$\;\;$ (normalized by \# of words)\\
%$\;\;$ Ordinal, Percent, Money, Set, Misc  N.E. \\
\hline
\textbf{Query Features}\\
$\cdot$ \% of query words coverage/matches\\
$\cdot$ Total event-type synonyms/hypernyms/ \\
$\;\;$ hyponyms coverage/matches.\\
%$\cdot$ Total event-type synonyms/hypernyms/ \\
%$\;\;$ hyponyms matches.\\
\hline
\textbf{Language Model Features}\\
$\cdot$ Avg. token log probability (domain lang. \\
$\;\;$ model)\\
$\cdot$ Avg. token log probability (background \\
$\;\;$ lang. model)\\
\hline
\textbf{Geo-tag Features}\\
$\cdot$ Median document distance to nearest  \\
$\;\;$ location cluster (current hour/prev. hour).\\
%$\cdot$ Median document distance to nearest  \\
%$\;\;$ location clusters (previous hour).\\
$\cdot$ Distance of first location in doc. to nearest \\
$\;\;$ location cluster (current hour/prev. hour).\\
%$\cdot$ Distance of first location in doc. to nearest \\
%$\;\;$ location cluster (previous hour).\\
\hline
\textbf{Temporal Features}\\
$\cdot$ Avg. tf-idf at current time.\\
$\cdot$ Change in avg. tf-idf since previous hour \\
$\;\;$  (up to 24 hours).\\
$\cdot$ Time since query/event start.\\
\hline
\end{tabular}
\caption{Salience model features.}
\end{figure}