### 2015-02-27 results.tex

\subsection{ROUGE}
\input{rouge_cum.tex}

Table~\ref{tab:rouge} shows our results for the retrospective evaluation using ROUGE. 
AP+Salience shows improvement on ngram precision and recall over the 
baselines. This improvement is statistically significant for all ngram and longest common substring
precision, recall, and F-measure
at the $\alpha = .01$ level
using the Wilcoxon signed-rank test.

AP+Salience maintains its performance above the baselines over time as well. Figure~\ref{fig:trouge}
shows the ROUGE-1 scores over time. 
We emphasize the difference in unigram precision (bigram precision is not shown but it follows
similar curve).
Within the initial days of the event, AP+Salience is able to take the lead over the over 
systems in ngram precision. 
The AP+Salience model is better able to find salient updates
earlier on; for the disaster domain, this is an especially important quality of the model.

Moreover, the AP+Salience's recall is not diminished by the high precision and remains competitive with AP.
Over time AP+Salience's recall also begins to pull away, while the other models start to suffer
from topic drift.

\subsection{Expected Gain and Comprehensiveness}

Figure~\ref{fig:nperf} shows the expected gain across a range of 
similarity thresholds, where thresholds closer to 1 are more conservative
estimates. The ranking of the systems remains constant across the 
sweep with 
AP+Salience beating all baseline systems.
Predicting salience in general is helpful for keeping a summary on topic as
the RS approach out performs the clustering only approaches
on expected gain.

When looking at the comprehensiveness of the summaries AP outperforms
 AP+Salience. The compromise encoded in the AP+Salience objective
function, between being representitive and being salient, is seen clearly here
where the performance of the AP+Salience methods is lower bounded by 
the salience focused RS
system and upper bounded by the clustering only AP system.
Overall, AP+Salience achieves the best balance of these two metrics.

\subsection{Feature Ablation}
\input{rouge_fa.tex}

Table~\ref{tab:farouge} shows the results of our feature ablation
tests. Removing the language models yields a statistically 
significant drop in both ngram recall and F-measure. 
Interestingly, removing the basic features leads to an
increase in both unigram and bigram precision; in the bigram
case this is enough to cause a statistically significant increase
in F-measure over the full model. In other words, the generic features
actually lead to an inferior model when we can incorporate more appropriate
domain specific features. \ckcomment{Kathy: maybe we should mention Spark Jones here about the idea that there is no such thing as generic summarization?}

\begin{figure}
    \includegraphics[]{rouge-time.eps}
\caption{System ROUGE-1 performance over time.}
\label{fig:trouge}
\end{figure}

\begin{figure}
  \includegraphics[]{nuggets-metrics.eps}
\caption{Avg. Gain and Comprehensiveness performance.}
\label{fig:nperf}
\end{figure}