### 2015-02-28 results.tex

\subsection{ROUGE}
\input{rouge_cum.tex}

Table~\ref{tab:rouge} shows our results for system output samples against the full summary of nuggets using ROUGE. 
%shows our results for the evaluation using ROUGE. 
%AP+Salience shows improvement on ngram precision and recall over the 
%baselines. 
This improvement is statistically significant for all ngram and longest common substring
precision, recall, and F-measure
at the $\alpha = .01$ level
using the Wilcoxon signed-rank test.

\begin{figure}
    \includegraphics[]{rouge-time.eps}
\caption{System ROUGE-1 performance over time.}
\label{fig:trouge}
\end{figure}

AP+Salience maintains its performance above the baselines over time as well. Figure~\ref{fig:trouge}
shows the ROUGE-1 scores over time. 
We show the difference in unigram precision (bigram precision is not shown but it follows
similar curve).
Within the initial days of the event, AP+Salience is able to take the lead over the over 
systems in ngram precision. 
The AP+Salience model is better able to find salient updates
earlier on; for the disaster domain, this is an especially important quality of the model.

Moreover, the AP+Salience's recall is not diminished by the high precision and remains competitive with AP.
Over time AP+Salience's recall also begins to pull away, while the other models start to suffer
from topic drift.

\subsection{Expected Gain and Comprehensiveness}
\begin{figure}
  \includegraphics[]{nuggets-metrics.eps}
\caption{Expected Gain and Comprehensiveness performance.}
\label{fig:nperf}
\end{figure}

Figure~\ref{fig:nperf} shows the expected gain across a range of 
similarity thresholds, where thresholds closer to 1 are more conservative
estimates. The ranking of the systems remains constant across the 
sweep with 
AP+Salience beating all baseline systems.
Predicting salience in general is helpful for keeping a summary on topic as
the RS approach out performs the clustering only approaches
on expected gain.

When looking at the comprehensiveness of the summaries AP outperforms
 AP+Salience. The compromise encoded in the AP+Salience objective
function, between being representitive and being salient, is seen clearly here
where the performance of the AP+Salience methods is lower bounded by 
the salience focused RS
system and upper bounded by the clustering only AP system.
Overall, AP+Salience achieves the best balance of these two metrics.

\subsection{Feature Ablation}
\input{rouge_fa.tex}

Table~\ref{tab:farouge} shows the results of our feature ablation
tests. Removing the language models yields a statistically 
significant drop in both ngram recall and F-measure. 
Interestingly, removing the basic features leads to an
increase in both unigram and bigram precision; in the bigram
case this is enough to cause a statistically significant increase
in F-measure over the full model. In other words, the generic features
actually lead to an inferior model when we can incorporate more appropriate
domain specific features.
This result mirrors Sparck Jones' claim that summarization should be in service of a purpose~\cite{?}.

Removing the language model and geographic relevance features leads to a
statistically significant drop in ROUGE-1 F1 scores. Unfortunately,
this is not the case for the temporal relevance features. We surmise that
these features are too strongly correlated with each other, 
i.e. the differences in TFIDF between hours are definitely not IID variables.