### 2015-02-28 conclusion.tex
ck rev

In this paper, we have presented an update summarization system for the disaster domain,
and demonstrated improved system performance by integrating sentence salience with clustering.

Addtionally, we have shown that feature groups specifically targeted for the domain of disaster yield
better summaries; specifically features that capture the language and location
of the domain.

In the future we would like to explore the application of the AP+Salience
 model and features
%KM3 - I think not just for more general domains (as it says something wrong about this one) but to a wider class of events.
to a wider class of events.

We would also like to further analyze the failure 
of the temporal relevance features; intuitively we believe time to be important
for the update summarization task.

%Addtionally we have presented several interesting feature groups for use in disaster summarization.
%In the future we would like to explore the applicability of this model and features
%to more general domains.