### 2015-05-29 conclusion.tex
spellinggit add acl2015.tex conclusion.tex dataset.tex experiments.tex introduction.tex methods.tex relatedwork.tex results.tex rouge_cum.tex nuggets-metrics-eps-converted-to.pdf rouge-time-eps-converted-to.pdf !

In this paper, we have presented an update summarization system for the 
disaster domain, and demonstrated improved system performance by integrating 
sentence salience with clustering.

We also have shown that features specifically targeted to the domain of 
disaster yield better summaries. We developed novel features that capture the 
language typical of different event types and that identify sentences specific
to the particular disaster based on location.

In the future we would like to explore the application of the 
\textsc{AP+Salience}
model and features to a wider class of events.