### 2015-05-29 dataset.tex

For the document stream, we use the news portion of the 2014 TREC KBA Stream 
Corpus \cite{frank2012building}. The documents from this corpus come from 
hourly crawls of the web covering October 2011 through February 2013.

Our experiments also make use of the TREC Temporal Summarization (TS) Track
 data from 2013 and 2014 \cite{aslam2013trec}.  This data includes 25 events 
and event metadata (e.g., a user search query for the event, the event type, 
and event evaluation time frame). All events occurred during the time span of 
the TREC KBA Stream Corpus. For each event we create a stream of relevant 
documents by selecting only documents that contain the complete set of query 
words.

Along with the metadata, NIST assessors constructed a set of ground truth 
nuggets for each event. Nuggets are brief and important text snippets that 
represent sub-events that should be conveyed by an ideal update summary. 
In order to accomplish this, for each event, assessors were provided with the
revision history of the Wikipedia page associated with the event. For example,
the revision history for the Wikipedia page for `Hurricane Sandy' will contain
text additions including those related to individual nuggets.  The assessment
task involves reviewing the Wikipedia revisions in the evaluation time frame 
and marking the text additions capturing a new, unique nugget.  More detail
on this process can be found in the track description \cite{aslam2013trec}.