### 2015-05-29 experiments.tex

We evaluate our system on two metrics: ROUGE \cite{lin2004rouge}, an automatic
summarization method and an evaluation of system expected gain and 
comprehensiveness---metrics adapted from the TREC TS track 
\cite{aslam2013trec}.

\subsection{Training and Testing}

Of the 25 events in the TREC TS data, 24 are covered by the news portion of 
the TREC KBA Stream Corpus. From these 24, we set aside three events to use as
a development set. All system salience and similarity threshold parameters are
tuned on the development set to maximize ROUGE-2 F1 scores.

We train a salience model for each event using 1000 sentences randomly sampled
from the event's document stream.

We perform a leave-one-out evaluation of each event. At test time, we predict 
a sentence's salience using the average predictions of the 23 other models.

\subsection{ROUGE Evaluation}

ROUGE measures the ngram overlap between a model summary and an automatically 
generated system summary. Model summaries for each event were constructed by 
concatenating the event's nuggets. Generally, ROUGE evaluation assumes both 
model and system summaries are of a bounded length. Since our systems are 
summarizing events over a span of two weeks time, the total length of our 
system output is much longer than the model. To address this, for each 
system/event pair, we sample with replacement 1000 random summaries of length 
less than or equal to the model summary (truncating the last sentence when 
neccessary). The final ROUGE scores for the system are the average scores 
from these 1000 samples.

Because we are interested in system performance over time, we also evaluate 
systems at 12 hour intervals using the same regime as above. The model 
summaries in this case are retrospective, and this evaluation reveals how 
quickly systems can cover  information in the model.

\subsection{Expected Gain and Comprehensiveness}

NIST developed metrics for evaluating update summarization systems as part of 
the TREC TS track. We present results on two of these metrics, the expected 
gain and comprehensiveness.

\textbf{Expected Gain } We treat the event's nuggets as unique units of 
information. When a system adds an update to its summary, it is potentially 
adding some of this nugget information. It would be instructive to know how 
much unique and novel information each update is adding on average to the 
summary. To that end, we define
\[ \mathrm{\mathbb{E}[Gain]} = \frac{|S_n|}{|S|}
\] 
where $S$ is the set of system updates, $S_n$ is the set of nuggets contained 
in $S$, and $|\cdot|$ is the number of elements in the set. To compute the 
set $S_n$ we match each system update to 0 or more nuggets, where an update 
matches a nugget if their semantic similarity is above a threshold. $S_n$ 
results from the unique set of nuggets matched. Because an update can map to 
more than one nugget, it is possible to receive an expected gain greater 
than 1. An expected gain of 1 would indicate that every sentence was both 
relevant and contained a unique piece of information.

\textbf{Comprehensiveness } Additionally, we can use the nuggets to measure 
the completeness of an update summary. We define
\[ \mathrm{Comprehensiveness} = \frac{|S_n|}{|N|}\]
where $N$ is the set of event nuggets. A comprehensiveness of 1 indicates that
the summary has covered all nugget information for the event; the maximum
attainable comprehensiveness is 1.

Update-nugget matches are computed automatically; a match exists if the 
semantic similarity of the update/nugget pair is above a threshold. 
Determining an optimal threshold to count matches is difficult so we evaluate 
at threshold values ranging from .5 to 1, where values closer to 1 are more 
conservative estimates of performance. A manual inspection of matches suggests
that semantic similarity values around .7 produce reasonable matches.
The average semantic similarity of manual matches performed by NIST assessors 
was much lower at approximately .25, increasing our confidence in the 
automatic matches in the .5--1 range.

\subsection{Model Comparisons}

We refer to our complete model as \textsc{AP+Salience}.  We compare this model
against several variants and baselines intended to measure the contribution of
different components. All thresholds for all runs are tuned on the development
set.

\textbf{Affinity Propagation only (\textsc{AP}) } The purpose of this model 
is to 
directly measure the effect of integrating salience and clustering by 
providing a baseline that uses the identical clustering component, but without
the salience information. In this model, input sentences are \textit{apriori}
equally likely to be exemplars; the salience values are uniformly set as the 
median value of the  input similarity scores, as is commonly used in the 
\textsc{AP} 
literature \cite{frey2007clustering}. After clustering a sentence batch, the 
exemplars are examined in order of increasing time since event start and 
selected as updates if their maximum similarity to the previous updates is 
less than $\lambda_{sim}$, as in the novelty filtering stage of 
\textsc{AP+Salience}.

\textbf{Hierarchichal Agglomerative Clustering (HAC)} We provide another 
clustering baseline, single-linkage hierarchichal agglomerative clustering.
We include this baseline to show that \textsc{AP+Salience} is not just an 
improvement
over \textsc{AP} but centrality driven methods in general. \textsc{HAC} was 
chosen over other 
clustering approaches because the number of clusters is not an explicit 
hyper-parameter. To produce flat clusters from the hierarchical clustering, we
flatten the \textsc{HAC} dendrogram using the cophenetic distance criteria, 
i.e. 
observations in each flat cluster have no greater a cophenetic distance than a
threshold. Cluster centers are determined to be the sentence with highest 
cosine similariy to the flat cluster mean. Cluster centers are examined in 
time order and are added to the summary if their similarity to previous 
updates is below a similarity threshold $\lambda_{sim}$ as is done in the 
\textsc{AP} model.

\textbf{Rank by Salience (RS)} 
We also isolate the impact of our salience model in order to demonstrate 
that the fusion of clustering and salience prediction improves over
predicting salience alone. In this model we predict the salience of sentences 
as in step 1 for \textsc{AP+Salience}. We omit the clustering phase (step 2).
Updates are selected identically to step 3 of \textsc{AP+Salience}, 
proceeding in order
of decreasing salience, selecting updates that are above a salience 
threshold $\lambda_{sal}$ and below a similarity threshold $\lambda_{sim}$
with respect to the previously selected updates.