### 2015-05-29 introduction.tex

\label{sec:introduction}

During crises, information is critical for first responders, crisis management
organizations, and those caught in the event. When the event is significant, 
as in the case of Hurricane Sandy, the amount of content produced by 
traditional news outlets, government agencies, relief organizations, and 
social media can vastly overwhelm those trying to monitor the situation. 
Crisis informatics \cite{palen2010vision} is dedicated to finding methods for 
sharing the right information in a timely fashion during such an event.
Research in this field has focused on human-in-the-loop approaches ranging 
from on the ground information gathering to crowdsourced reporting and 
disaster management \cite{starbird2013working}.

Multi-document summarization has the potential to assist the crisis 
informatics community. Automatic summarization could deliver relevant and 
salient information at regular intervals, even when human volunteers are 
unable to. Perhaps more importantly it could help filter out unnecessary and 
irrelevant detail when the volume of incoming information is large. While 
methods for identifying, tracking, and summarizing events from text based 
input have been explored extensively
\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}, 
these experiments were not developed to handle streaming data from a
heterogeneous environment at web scale. These methods also rely heavily on 
redundancy which is suboptimal for time sensitive domains where there is a 
high cost in delaying information.

In this paper, we present an update summarization system to track events
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster, and integrates these predictions into
a clustering based multi-document summarization system. We demonstrate that 
combining salience with clustering produces more relevant summaries compared 
to baselines using clustering or relevance alone.  Our experiments suggest 
that this is because our system is better able to adapt to dynamic changes in 
input volume that adversely affect methods that use redundancy as a proxy for 
salience.

In addition to the tight integration between clustering and salience
prediction, our approach also exploits knowledge about the event to determine
salience. Thus, salience represents both how typical a sentence is of the  
event
type (e.g., industrial accident, hurricane, riot) and whether it specifies 
information
about this particular event. 
Our feature representation includes a set of language models, one for each
event type, to measure the typicality of the sentence with regard to the 
current event, the distance of mentioned locations from the center of
the event, and the change in word frequencies over the time of the event.
While we evaluate these features in the domain of disasters, this approach is 
generally applicable to many update summarization tasks.

Our approach achieves a statistically significant improvement in ROUGE scores 
compared to multiple baselines. Additionally, we introduce novel methods for 
estimating the average information gain each update provides and how 
completely the update summary covers the event it is tracking; our system's 
updates contain more relevant information on average than the competing 
baselines.

The remainder of the paper is organized as follows. We begin with a review of 
related work in the information retrieval and multi-document summarization 
literature. Section~\ref{sec:methods} outlines the details of our salience 
and summarization models. Next we describe our data (Section~\ref{sec:data}) 
and experiments (Section~\ref{sec:exper}). Finally, we discuss our results 
(Section~\ref{sec:results}) and conclude the paper.