### 2015-05-29 results.tex

\subsection{ROUGE}
\input{rouge_cum.tex}

Table~\ref{tab:rouge} shows our results for system output samples against the 
full summary of nuggets using ROUGE. This improvement is statistically 
significant for all ngram precision, recall, and F-measures at the 
$\alpha = .01$ level using the Wilcoxon signed-rank test.

\begin{figure}
    \includegraphics[]{rouge-time.eps}
\caption{System ROUGE-1 performance over time.}
\label{fig:trouge}
%\vspace{-14pt}
\end{figure}

\textsc{AP+Salience} maintains its performance above the baselines over time 
as well. Figure~\ref{fig:trouge} shows the ROUGE-1 scores over time. We show 
the difference in unigram precision (bigram precision is not shown but it 
follows similar curve). Within the initial days of the event, 
\textsc{AP+Salience} is able to take the lead over the over systems in ngram 
precision. The \textsc{AP+Salience} model is better able to find salient 
updates earlier on; for the disaster domain, this is an especially important 
quality of the model.

Moreover, the \textsc{AP+Salience}'s recall is not diminished by the high 
precision and remains competitive with \textsc{AP}. Over time 
\textsc{AP+Salience}'s recall also begins to pull away, while the other models
start to suffer from topic drift.

\subsection{Expected Gain and Comprehensiveness}
\begin{figure}[h]
  \includegraphics[]{nuggets-metrics.eps}
\caption{Expected Gain and Comprehensiveness performance.}
\label{fig:nperf}
%\vspace{-12pt}
\end{figure}

Figure~\ref{fig:nperf} shows the expected gain across a range of similarity 
thresholds, where thresholds closer to 1 are more conservative estimates. 
The ranking of the systems remains constant across the sweep with 
\textsc{AP+Salience} beating all baseline systems. Predicting salience in 
general is helpful for keeping a summary on topic as the \textsc{RS} approach 
out performs the clustering only approaches on expected gain.

When looking at the comprehensiveness of the summaries \textsc{AP} outperforms
\textsc{AP+Salience}. The compromise encoded in the \textsc{AP+Salience} 
objective function, between being representative and being salient, is seen 
clearly here where the performance of the \textsc{AP+Salience} methods is 
lower bounded by the salience focused \textsc{RS} system and upper bounded by 
the clustering only \textsc{AP} system. Overall, \textsc{AP+Salience} achieves
the best balance of these two metrics.

\subsection{Feature Ablation}
\input{rouge_fa.tex}

Table~\ref{tab:farouge} shows the results of our feature ablation tests. 
Removing the language models yields a statistically significant drop in both 
ngram recall and F-measure. Interestingly, removing the basic features leads 
to an increase in both unigram and bigram precision; in the bigram case this 
is enough to cause a statistically significant increase in F-measure over the 
full model. In other words, the generic features actually lead to an inferior 
model when we can incorporate more appropriate domain specific features.
The result mirrors Sparck Jones' claim that generic approaches to 
summarization cannot produce a useful summary \cite{ksj98}.

Removing the language model and geographic relevance features leads to a
statistically significant drop in ROUGE-1 F1 scores. Unfortunately,
this is not the case for the temporal relevance features. We surmise that
these features are too strongly correlated with each other, 
i.e. the differences in TF*IDF between hours are definitely not i.i.d. 
variables.