### 2015-05-29 outputs.tex
Added example outputs.

\begin{figure*}
\begin{tabular}{|l m{15cm}|}
\multicolumn{2}{c}{\textbf{Hurricane Sandy}}\\
\hline
\hline
%\small
%\tabitem The forecast map shows Sandy reaching eastern Cuba by early Thursday 
%    before heading to the Bahamas.\\
%\small
%\tabitem Jamaica's government issued a hurricane warning on Tuesday morning and 
%    announced schools would be closed on Wednesday.\\
\small
\tabitem  & \small Dangerous flash floods and mudslides set off by Sandy were a threat for the
   island of roughly 2.7 million inhabitants, Jamaica's meteorological service
   said.\\
\small
\tabitem  & \small A few reliable forecast models bring Sandy close enough to the coast to 
   produce heavy rains, strong winds and beach erosion.\\
\small
\tabitem  & \small Max winds are 65 mph with strengthening to a hurricane expected in the next
   12 to 18 hours.\\
\small
\tabitem & \small The two international airports, as well as schools and businesses, 
         will remain closed today until the hurricane warning, which is now in
         effect for the island, is lifted. \\
\hline
%at 11 a.m., sandy was in the caribbean about 65 miles south of kingston , jamaica , moving north at 13 mph with sustained winds of 80 mph.
%"""""""we heard on the radio that the hurricane was coming this way,"""""""" he said in the poor kingston community of standpipe, situated next to one of the debris-clogged gullies that crisscross the capital."""""""
\end{tabular}

~\\
~\\
\begin{tabular}{|l m{15cm}|}
\multicolumn{2}{c}{\textbf{2012 Pakistan Garment Factory Fires}}\\
\hline
\hline
%\small
%\tabitem Medical officials said that some had died of suffocation while others
%         were burnt alive as the fire took hold. \\
\small
\tabitem & \small The fire broke out when people in the building were trying to start 
         their generator after the electricity went out. \\
\small
\tabitem & \small Pakistani television showed pictures of what appeared to be a 
         three-story building with flames leaping from the top-floor windows 
         and smoke billowing into the night sky. \\
\small
\tabitem & \small The people went to the back side of the building but there was no 
         access, so we had to made forceful entries and rescue the people, 
         said Numan Noor, a firefighter on the scene. \\
\small
\tabitem & \small ``We have recovered 63 bodies, including three found when we reached 
         the basement of the building,'' Karachi fire chief Ehtesham Salim told
         AFP on Wednesday. \\
%\small
%\tabitem Salim added that the blaze was Karachi’s ``biggest fire in terms of 
%         deaths in decades.'' \\
%\small
%\tabitem The garment trade as a whole is vital to Pakistan’s shaky economy. \\
\hline
\end{tabular}

~\\
~\\
\begin{tabular}{|l m{15cm}|}
\multicolumn{2}{c}{\textbf{2012 Romanian Protests}}\\
\hline
\hline
\small
\tabitem & \small Clashes between riot police and demonstrators have also erupted in 
         the Romanian capital Bucharest for a third day in a row. \\
\small
\tabitem & \small BOC urged Romanians to understand that tough austerity measures are 
         needed to avoid a default. \\
\small
\tabitem & \small More than 1,000 protesters rallied in Bucharest's main university 
         square, blocking traffic. \\
\small
\tabitem & \small Bucharest : a Romanian medical official says 59 people suffered 
         injuries as days of protests against the government and austerity 
         measures turned violent. \\
%\small
%\tabitem Most Romanians agree that only fundamental reform can save the 
%         country's ailing and corrupt system. \\
%\small
%\tabitem Many Romanians feared this would only increase corruption and create 
%         a further divide between the classes, leading to a two-tier system in
%         which only the wealthy would be able to afford care. \\
\hline
\end{tabular}

\caption{\textsc{AP+Salience} summary excerpts.}
\label{fig:summaries}
\end{figure*}