### 2015-05-29 outputs.tex
Added event type examples.

\begin{figure*}
%\begin{tabular}{|l m{15cm}|}
%\multicolumn{2}{c}{\textbf{Hurricane Sandy}}\\
%\hline
%\hline
%%\small
%%\tabitem The forecast map shows Sandy reaching eastern Cuba by early Thursday 
%%    before heading to the Bahamas.\\
%%\small
%%\tabitem Jamaica's government issued a hurricane warning on Tuesday morning and 
%%    announced schools would be closed on Wednesday.\\
%\small
%\tabitem  & \small Dangerous flash floods and mudslides set off by Sandy were a threat for the
%   island of roughly 2.7 million inhabitants, Jamaica's meteorological service
%   said.\\
%\small
%\tabitem  & \small A few reliable forecast models bring Sandy close enough to the coast to 
%   produce heavy rains, strong winds and beach erosion.\\
%\small
%\tabitem  & \small Max winds are 65 mph with strengthening to a hurricane expected in the next
%   12 to 18 hours.\\
%\small
%\tabitem & \small The two international airports, as well as schools and businesses, 
%         will remain closed today until the hurricane warning, which is now in
%         effect for the island, is lifted. \\
%\hline
%%at 11 a.m., sandy was in the caribbean about 65 miles south of kingston , jamaica , moving north at 13 mph with sustained winds of 80 mph.
%%"""""""we heard on the radio that the hurricane was coming this way,"""""""" he said in the poor kingston community of standpipe, situated next to one of the debris-clogged gullies that crisscross the capital."""""""
%\end{tabular}
%
%~\\
%~\\
\begin{tabular}{|l m{15cm}|}
\multicolumn{2}{c}{\textbf{2012 Pakistan Garment Factory Fires}}\\
\hline
\hline
%\small
%\tabitem Medical officials said that some had died of suffocation while others
%         were burnt alive as the fire took hold. \\
\small
\tabitem & \small The fire broke out when people in the building were trying to start 
         their generator after the electricity went out. \\
\small
\tabitem & \small Pakistani television showed pictures of what appeared to be a 
         three-story building with flames leaping from the top-floor windows 
         and smoke billowing into the night sky. \\
\small
\tabitem & \small The people went to the back side of the building but there was no 
         access, so we had to made forceful entries and rescue the people, 
         said Numan Noor, a firefighter on the scene. \\
\small
\tabitem & \small ``We have recovered 63 bodies, including three found when we reached 
         the basement of the building,'' Karachi fire chief Ehtesham Salim told
         AFP on Wednesday. \\
%\small
%\tabitem Salim added that the blaze was Karachi’s ``biggest fire in terms of 
%         deaths in decades.'' \\
%\small
%\tabitem The garment trade as a whole is vital to Pakistan’s shaky economy. \\
\hline
\end{tabular}

~\\
~\\
\begin{tabular}{|l m{15cm}|}
\multicolumn{2}{c}{\textbf{2012 Romanian Protests}}\\
\hline
\hline
\small
\tabitem & \small Clashes between riot police and demonstrators have also erupted in 
         the Romanian capital Bucharest for a third day in a row. \\
\small
\tabitem & \small BOC urged Romanians to understand that tough austerity measures are 
         needed to avoid a default. \\
\small
\tabitem & \small More than 1,000 protesters rallied in Bucharest's main university 
         square, blocking traffic. \\
\small
\tabitem & \small Bucharest : a Romanian medical official says 59 people suffered 
         injuries as days of protests against the government and austerity 
         measures turned violent. \\
%\small
%\tabitem Most Romanians agree that only fundamental reform can save the 
%         country's ailing and corrupt system. \\
%\small
%\tabitem Many Romanians feared this would only increase corruption and create 
%         a further divide between the classes, leading to a two-tier system in
%         which only the wealthy would be able to afford care. \\
\hline
\end{tabular}

\caption{\textsc{AP+Salience} summary excerpts.}
\label{fig:summaries}
\end{figure*}